quollr: An R Package for Visualizing 2-D Models from Non-linear Dimension Reductions in High Dimensional Space

### 2025-05-15 paper-quollr.Rmd
Added Usage section

<!-- ## Usage -->

<!-- add about main function to gen 2D and highD models-->
<!--Discuss the model can be generated with bin centroids or bin means-->

@@

## Main function

@@

## Constructing the $2\text{-}D$ Model

@@

### Scaling the Data

@@

### Computing hexagon grid configurations

@@

### Binning the data

@@

#### Generating all possible centroids in a hexagonal grid

@@

#### Creating the coordinates of the hexagons

@@

#### Assigning data points to their respective hexagons

@@

#### Computing the standardized number of points within each hexagon

@@

#### Mapping the points to their corresponding hexagonal bins

@@

### Obtaining bin centroids

@@

### Indicating neighbors by line segments connecting centroids

@@

### Identifying and removing low-density hexagons

@@

## Lifting the model into high dimensions

@@

## Prediction

@@

## Compute residuals and Mean Square Error (MSE)

As a Goodness of fit statistics for the model, `glance()` is used to compute residuals and MSE. These metrics are used to assess how well the fitted model will capture the underlying structure of the $p\text{-}D$ data.

@@

From these per-dimension errors, the function computes absolute error which is the sum of absolute differences across all dimensions and observations and the Mean squared error (MSE) which is the average of the total squared error per point.

These metrics are returned in a tibble as `Error` (absolute error) and `MSE` (mean squared error).

@@

## Visualizations

@@

### Hexagonal grid

@@

### Triangular mesh

@@

### $p\text{-}D$ model visualization

@@

### Link plots

>>>>

## Usage

@@

### Main function

@@

### Constructing the $2\text{-}D$ Model

@@

#### Scaling the Data

@@

#### Computing hexagon grid configurations

@@

#### Binning the data

@@

##### Generating all possible centroids in a hexagonal grid

@@

##### Creating the coordinates of the hexagons

@@

##### Assigning data points to their respective hexagons

@@

##### Computing the standardized number of points within each hexagon

@@

##### Mapping the points to their corresponding hexagonal bins

@@

#### Obtaining bin centroids

@@

#### Indicating neighbors by line segments connecting centroids

@@

#### Identifying and removing low-density hexagons

@@

### Lifting the model into high dimensions

@@

### Prediction

@@

### Compute residuals and Root Mean Square Error (RMSE)

As a Goodness of fit statistics for the model, `glance()` is used to compute residuals and RMSE. These metrics are used to assess how well the fitted model will capture the underlying structure of the $p\text{-}D$ data.

@@

From these per-dimension errors, the function computes absolute error which is the sum of absolute differences across all dimensions and observations and the RMSE which is the average of the total squared error per point.

These metrics are returned in a tibble as `Error` (absolute error) and `RMSE` (root mean squared error).

@@

### Visualizations

@@

#### Hexagonal grid

@@

#### Triangular mesh

@@

#### $p\text{-}D$ model visualization

@@

#### Link plots

### 2025-05-15 paper-quollr.tex
Added Usage section

@@

\subsection{Main function}\label{main-function}

@@

\subsection{\texorpdfstring{Constructing the \(2\text{-}D\) Model}{Constructing the 2\textbackslash text\{-\}D Model}}\label{constructing-the-2text-d-model}

@@

\subsubsection{Scaling the Data}\label{scaling-the-data}

@@

\subsubsection{Computing hexagon grid configurations}\label{computing-hexagon-grid-configurations}

@@

\subsubsection{Binning the data}\label{binning-the-data}

@@

\paragraph{Generating all possible centroids in a hexagonal grid}\label{generating-all-possible-centroids-in-a-hexagonal-grid}

@@

\paragraph{Creating the coordinates of the hexagons}\label{creating-the-coordinates-of-the-hexagons}

@@

\paragraph{Assigning data points to their respective hexagons}\label{assigning-data-points-to-their-respective-hexagons}

@@

\paragraph{Computing the standardized number of points within each hexagon}\label{computing-the-standardized-number-of-points-within-each-hexagon}

@@

\paragraph{Mapping the points to their corresponding hexagonal bins}\label{mapping-the-points-to-their-corresponding-hexagonal-bins}

@@

\subsubsection{Obtaining bin centroids}\label{obtaining-bin-centroids}

@@

\subsubsection{Indicating neighbors by line segments connecting centroids}\label{indicating-neighbors-by-line-segments-connecting-centroids}

@@

\subsubsection{Identifying and removing low-density hexagons}\label{identifying-and-removing-low-density-hexagons}

@@

\subsection{Lifting the model into high dimensions}\label{lifting-the-model-into-high-dimensions}

@@

\subsection{Prediction}\label{prediction}

@@

\subsection{Compute residuals and Mean Square Error (MSE)}\label{compute-residuals-and-mean-square-error-mse}

As a Goodness of fit statistics for the model, \texttt{glance()} is used to compute residuals and MSE. These metrics are used to assess how well the fitted model will capture the underlying structure of the \(p\text{-}D\) data.

@@

From these per-dimension errors, the function computes absolute error which is the sum of absolute differences across all dimensions and observations and the Mean squared error (MSE) which is the average of the total squared error per point.

These metrics are returned in a tibble as \texttt{Error} (absolute error) and \texttt{MSE} (mean squared error).

@@

\subsection{Visualizations}\label{visualizations}

@@

\subsubsection{Hexagonal grid}\label{hexagonal-grid}

@@

\subsubsection{Triangular mesh}\label{triangular-mesh}

@@

\subsubsection{\texorpdfstring{\(p\text{-}D\) model visualization}{p\textbackslash text\{-\}D model visualization}}\label{ptext-d-model-visualization}

@@

\subsubsection{Link plots}\label{link-plots}

>>>>

\subsection{Usage}\label{usage}

@@

\subsubsection{Main function}\label{main-function}

@@

\subsubsection{\texorpdfstring{Constructing the \(2\text{-}D\) Model}{Constructing the 2\textbackslash text\{-\}D Model}}\label{constructing-the-2text-d-model}

@@

\paragraph{Scaling the Data}\label{scaling-the-data}

@@

\paragraph{Computing hexagon grid configurations}\label{computing-hexagon-grid-configurations}

@@

\paragraph{Binning the data}\label{binning-the-data}

@@

\subparagraph{Generating all possible centroids in a hexagonal grid}\label{generating-all-possible-centroids-in-a-hexagonal-grid}

@@

\subparagraph{Creating the coordinates of the hexagons}\label{creating-the-coordinates-of-the-hexagons}

@@

\subparagraph{Assigning data points to their respective hexagons}\label{assigning-data-points-to-their-respective-hexagons}

@@

\subparagraph{Computing the standardized number of points within each hexagon}\label{computing-the-standardized-number-of-points-within-each-hexagon}

@@

\subparagraph{Mapping the points to their corresponding hexagonal bins}\label{mapping-the-points-to-their-corresponding-hexagonal-bins}

@@

\paragraph{Obtaining bin centroids}\label{obtaining-bin-centroids}

@@

\paragraph{Indicating neighbors by line segments connecting centroids}\label{indicating-neighbors-by-line-segments-connecting-centroids}

@@

\paragraph{Identifying and removing low-density hexagons}\label{identifying-and-removing-low-density-hexagons}

@@

\subsubsection{Lifting the model into high dimensions}\label{lifting-the-model-into-high-dimensions}

@@

\subsubsection{Prediction}\label{prediction}

@@

\subsubsection{Compute residuals and Root Mean Square Error (RMSE)}\label{compute-residuals-and-root-mean-square-error-rmse}

As a Goodness of fit statistics for the model, \texttt{glance()} is used to compute residuals and RMSE. These metrics are used to assess how well the fitted model will capture the underlying structure of the \(p\text{-}D\) data.

@@

From these per-dimension errors, the function computes absolute error which is the sum of absolute differences across all dimensions and observations and the RMSE which is the average of the total squared error per point.

These metrics are returned in a tibble as \texttt{Error} (absolute error) and \texttt{RMSE} (root mean squared error).

@@

\subsubsection{Visualizations}\label{visualizations}

@@

\paragraph{Hexagonal grid}\label{hexagonal-grid}

@@

\paragraph{Triangular mesh}\label{triangular-mesh}

@@

\paragraph{\texorpdfstring{\(p\text{-}D\) model visualization}{p\textbackslash text\{-\}D model visualization}}\label{ptext-d-model-visualization}

@@

\paragraph{Link plots}\label{link-plots}
